---
title: "Variation on Scatterplots"
author: "Diardano Raihan"
date: "2022-07-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import the Libraries and Dataset

```{r dataset}
library(tidyverse)
library(knitr)
library(ggthemes)
library(RColorBrewer)
cces <- read_csv("cces_sample_coursera.csv")
cel <- read_csv("cel_dataset_coursera.csv")
```

# Variations on Scatter Plots
## Add a best-fit line with geom_smooth
Reference: 
- Adapted from https://r-graphics.org/recipe-scatter-fitlines
- Adapted from https://ggplot2.tidyverse.org/reference/geom_smooth.html

```{r plot1}
# Let's see the normal scatter plot look like
ggplot(cces, aes(x = educ, y = ideo5))+
  geom_jitter()

# Add a linear model "lm" that fit the plot
ggplot(cces, aes(x = educ, y = ideo5))+
  geom_jitter()+
  geom_smooth(method = "lm")

# Add a loess curve moder "lm" that fit the plot
ggplot(cces, aes(x = educ, y = ideo5))+
  geom_jitter()+
  geom_smooth(method = "loess")

# Change the confidence level 
ggplot(cces, aes(x = educ, y = ideo5))+
  geom_jitter()+
  geom_smooth(method = "lm", level = .9)
```

- Be very careful with changing the confidence level if you're not statistically inclined, because doing statistical inference with this kind of approach is pretty dicey. 

- Exploratory analysis is fine so if you want to draw some best fit lines to your scatter plot to get a sense for the data just sort of an intuition is an okay thing to do.

## Scatter Plot Matrix
- Another use of scatter plot for EDA is to use a scatter plot matrix which will show you the bivariate relationship several different variables in your data at once. 
```{r plot2}
# install.packages("GGally")
library(GGally)

# Make up some numeric data.
# Two variables will be positively correlated, and the third will be negatively correlated with the first.
var1 <- runif(100, 0, 1)
var2 <- var1 + rnorm(100, 1, .2)
var3 <- var1*(-rnorm(100, 1, .2))

df <- tibble(var1, var2, var3)
df

ggpairs(df)
```

## Customize the Matrix Figures by Custom Function
```{}

```