---
title: "Variation on Scatterplots"
author: "Diardano Raihan"
date: "2022-07-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import the Libraries and Dataset

```{r dataset}
library(tidyverse)
library(knitr)
library(ggthemes)
library(RColorBrewer)
cces <- read_csv("cces_sample_coursera.csv")
cel <- read_csv("cel_dataset_coursera.csv")
```

# Variations on Scatter Plots
## Add a best-fit line with geom_smooth
Reference: 
- Adapted from https://r-graphics.org/recipe-scatter-fitlines
- Adapted from https://ggplot2.tidyverse.org/reference/geom_smooth.html

```{r plot1}
# Let's see the normal scatter plot look like
ggplot(cces, aes(x = educ, y = ideo5))+
  geom_jitter()

# Add a linear model "lm" that fit the plot
ggplot(cces, aes(x = educ, y = ideo5))+
  geom_jitter()+
  geom_smooth(method = "lm")

# Add a loess curve moder "lm" that fit the plot
ggplot(cces, aes(x = educ, y = ideo5))+
  geom_jitter()+
  geom_smooth(method = "loess")

# Change the confidence level 
ggplot(cces, aes(x = educ, y = ideo5))+
  geom_jitter()+
  geom_smooth(method = "lm", level = .9)
```

- Be very careful with changing the confidence level if you're not statistically inclined, because doing statistical inference with this kind of approach is pretty dicey. 

- Exploratory analysis is fine so if you want to draw some best fit lines to your scatter plot to get a sense for the data just sort of an intuition is an okay thing to do.

## Scatter Plot Matrix
- Another use of scatter plot for EDA is to use a scatter plot matrix which will show you the bivariate relationship several different variables in your data at once. 
- We can achieve this by using `ggplot` function.
  - Reference: https://rkabacoff.github.io/datavis/Other.html#scatterplot-matrix
```{r plot2}
# install.packages("GGally")
library(GGally)

# Make up some numeric data.
# Two variables will be positively correlated, and the third will be negatively correlated with the first.
var1 <- runif(100, 0, 1)
var2 <- var1 + rnorm(100, 1, .2)
var3 <- var1*(-rnorm(100, 1, .2))

df <- tibble(var1, var2, var3)
df

# Let's plot the scatterplot matrix using ggpair()
ggpairs(df)
```

## Customize the Matrix Figures by Custom Function
```{r plot3}
# Write your own function for the scatterplot
my_scatter <- function(data, mapping){
  ggplot(data = data, mapping = mapping)+
    geom_jitter(color = "red")
}

# Write your own function for density plot
my_density <- function(data, mapping){
  ggplot(data = data, mapping = mapping)+
    geom_density(alpha = .05, 
                 fill = "red")
}

# Substitute your functions for the functions that ggpairs() uses to draw the figures
ggpairs(df, 
        lower = list(continuous = my_scatter),
        diag = list(continuous = my_density))
```

## Correlation Plots
- Reference: https://rkabacoff.github.io/datavis/Models.html#Corrplot
```{r plot4}
# install.packages("ggcorrplot")
library(ggcorrplot)
library(ggthemes)

# Let's use some example survey data
df <- cces %>% select("educ", "pid7", "pew_religimp")

# Calculate correlation coefficients
r <- cor(df, use = 'complete.obs')
?cor

# generate the correlation plot
ggcorrplot(r)

# show just the lower part of the figure (to avoid 1 correlations on the diagonal)
ggcorrplot(r, type = "lower")

# modify some visual elements
ggcorrplot(r, type = "lower", 
           title = "Correlations", 
           colors = c("yellow", "green", "blue"), 
           outline.color = "purple")

# use ggtheme is you want
ggcorrplot(r, type = "lower", 
           title = "Correlations")+
  theme_wsj()
```

## Cleveland Dot Plots
- Adapted from https://r-graphics.org/recipe-bar-graph-dot-plot

```{r plot5}
# Use some of the congress data
cel_114 <- cel %>% filter(congress == 114)

members <- sample_n(cel_114, 25)

# points only
ggplot(members, aes(x = les, y = thomas_name))+
  geom_point()

# some refinements
ggplot(members, aes(x = les, y = reorder(thomas_name, les)))+ # reorder in descending
  geom_point(size = 5)+
  theme(panel.grid.major.x = element_blank(), 
        panel.grid.minor.x = element_blank(), 
        panel.grid.major.y = element_line(linetype = "dashed", color = "blue"))+
  labs(x = "Legislative Effectiveness", y = "")
```

## Lollipop Figure
- Although it falls under the variations of scatterplot, it is also just the very refined simple way of making a bar plot essentially.
- https://www.r-graph-gallery.com/300-basic-lollipop-plot.html

```{r plot6}
ggplot(members, aes(x = reorder(thomas_name, les), y = les))+ # reorder in descending
  geom_point(size = 5)+
  geom_segment(aes(x = thomas_name, xend = thomas_name, y = 0, yend = les))+
  theme(axis.text.x = element_text(angle = 90))
```