---
title: "3_variations_on_flow_n_circle_diagrams"
author: "Diardano Raihan"
date: "2022-07-21"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Import the Libraries and Dataset

```{r dataset}
library(tidyverse)
library(knitr)
library(ggthemes)
library(RColorBrewer)
```

# Flows

## Alluvial Diagrams
- A good visualization to show parts of a whole world connections between different categories and data.
- Let's create some fake data about student performance in classes.
  - 21 students are divided in 3 groups, they are men and women, and they get High Pass, Pass, or Fail.
- Reference: https://github.com/corybrunson/ggalluvial

```{r data}
group1 <- tibble(groupid = groupid <- rep("group1", 7), 
                 studentID = sample(seq(from = 1, to = 20), 7),
                 gender = sample(c("M", "F"), 7, replace = TRUE), 
                 grades = sample(c("High Pass", "Pass", "Fail"), 7, replace = TRUE))

group2 <- tibble(groupid = groupid <- rep("group2", 7), 
                 studentID = sample(seq(from = 21, to = 30), 7),
                 gender = sample(c("M", "F"), 7, replace = TRUE), 
                 grades = sample(c("High Pass", "Pass", "Fail"), 7, replace = TRUE))

group3 <- tibble(groupid = groupid <- rep("group3", 7), 
                 studentID = sample(seq(from = 31, to = 40), 7),
                 gender = sample(c("M", "F"), 7, replace = TRUE), 
                 grades = sample(c("High Pass", "Pass", "Fail"), 7, replace = TRUE))

students <- bind_rows(group1, group2, group3)
students

# summarize data
students_table <- students %>%
  group_by(groupid, gender, grades) %>%
  count()
```

- Now, let's draw the alluvial framework

```{r plot1}
# install.packages("ggalluvial")
library(ggalluvial)

# barebones alluvial
ggplot(students_table, 
       aes(axis1 = groupid, 
           axis2 = grades, 
           y = n))+
  geom_alluvium()

# add strata
ggplot(students_table, 
       aes(axis1 = groupid, 
           axis2 = grades, 
           y = n))+
  geom_alluvium()+
  geom_stratum()

# add labels
ggplot(students_table, 
       aes(axis1 = groupid, 
           axis2 = grades, 
           y = n))+
  geom_alluvium()+
  geom_stratum()+
  geom_text(stat = "stratum", 
            aes(label = after_stat(stratum)))

# change colors
ggplot(students_table, 
       aes(axis1 = groupid, 
           axis2 = grades, 
           y = n))+
  geom_alluvium(aes(fill = gender))+
  geom_stratum()+
  geom_text(stat = "stratum", 
            aes(label = after_stat(stratum)))+
  scale_fill_manual(values = c("purple", "green"))
```