---
output:
  pdf_document: default
  html_document: default
---
# Introduction to `ggplot` 

- __gg__ in `ggplot` stands for the grammar of graphics.

- The recipe: 
  - Tidy data
  - Mapping _aesthetics_ to the plot
  - Geoms - geometric shapes on the plot
  - Coordinate system
  - Lables, guides (legends), and other annotations. 

- Dataset
  - Data from the Center for Effective Lawmaking (CEL) from Volden and Wisemen
  - This dataset is great for visualization purposes.
  - Review the codebook and look at the .csv attached.

- Business Problem
  - How does the number of bills passed by a member in a Congress relate to seniority? 
  - How long members have been in the chamber?
  - We'll focus on the 115th Congress.
  
# Load the libraries
```{r setup}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```

# Load the Data
```{r dataset, echo=TRUE}
data <- read_csv("cel_dataset_coursera.csv")

# see the column names
names(data)

# see the dimension of data
dim(data)

# see the statistics summary
summary(data$all_bills)
table(data$year)
```

# Data Visualization

## Basic ggplot

- To create a scatter plot, once we add the aesthetic to the `ggplot` function, we can add the geometric shape to it by adding a function called `geom_point`. 
- Note that we don't need to pass any parameter inside the function because it inherits the data and aesthetic mapping from the original from the original ggplot command.

```{r myFirstVisual}
# Filter the data we want
fig115 <- data %>% filter(congress == 115) %>% select("seniority", "all_pass")
head(fig115)

# set up the data and aesthetic.
# Aesthetic mapping connects visual elements of the figure with variables in the data.
ggplot(fig115, aes(x = seniority, y = all_pass)) # it will generate a blank plot with x and y axis

# add the marks or points
ggplot(fig115, aes(x = seniority, y = all_pass))+
  geom_point()

```

## Add Jitter and Labels
- Notice that many points in the scatterplot are on top of each other. 
- To solve this, we can use `geom_jitter` to add randoms noise to the data to avoid overplotting. 
- This technique is useful to maximize the reader's ability to interpret the data while minimizing any kind of deceptive or manipulating practice.

```{r mySecondVisual}
# jitter adds random noise tot he ata to avoid overplotting
ggplot(fig115, aes(x = seniority, y = all_pass))+
  geom_jitter()

# add some labels and a title
ggplot(fig115, aes(x = seniority, y = all_pass))+
  geom_jitter()+
  labs(x = "Seniority", y = "Bills Passed", title = "Seniority and Bills Passed in the 115th Congress")

```

## Add and Modify the Color

```{r myThirdVisual}
# modify the filter and select to grab "dem"
fig115 <- data %>% 
  filter(congress == 115) %>%
  select("seniority", "all_pass", "dem")

ggplot(fig115, aes(x = seniority, y = all_pass, color = dem))+
  geom_jitter()+
  labs(x = "Seniority", y = "Bills Passed", title = "Seniority and Bills Passed in the 115th Congress")

```

- Notice that the color legend is strange since it's not supposed to be a continuous one rather than a discrete one. Let's fix it!

```{r myFourthVisual}
party <- recode(fig115$dem, '1' = 'Democrat', '0' = 'Republican')

fig115 <- add_column(fig115, party)

ggplot(fig115, aes(x = seniority, y = all_pass, color = party))+
  geom_jitter()+
  labs(x = "Seniority", y = "Bills Passed", title = "Seniority and Bills Passed in the 115th Congress")
```

- The color that we just saw is set automatically by R. Let's modify it!

- let's make the colors match traditional blue democrats and red republicans.

```{r myFifthVisual}
ggplot(fig115, aes(x = seniority, y = all_pass, color = party))+
  geom_jitter()+
  labs(x = "Seniority", y = "Bills Passed", title = "Seniority and Bills Passed in the 115th Congress")+
  scale_color_manual(values = c("blue", "red"))
```

## Separate the Plot into Subplots
- Let's make the plot into two separate plots using `face_wrap`.
- Facetting is creating tables of graphics by splitting the data into subsets and displaying the same graph for each subset.
```{r myLastPlot}
ggplot(fig115, aes(x = seniority, y = all_pass, color = party))+
  geom_jitter()+
  labs(x = "Seniority", y = "Bills Passed", title = "Seniority and Bills Passed in the 115th Congress")+
  scale_color_manual(values = c("blue", "red"))+
  facet_wrap(~party)
```