---
title: "Distribution Plot"
output: pdf_document
date: "2022-07-17"
---

```{r setup, include=TRUE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```

# Load the Dataset

```{r dataset}
data <- read_csv('cces_sample_coursera.csv')
```

# Data Visualization

Remember the same process that we will do everytime we want to generate plot using `ggplot`: 
1. Start with the `ggplot` function. 
2. Add the `geom_` that you want. 
3. Mess around with the additional visual elements such as _title_, _lables_, etc.

## Boxplot
Reference:
  - https://r-graphics.org/recipe-quick-boxplot
  - https://r-graphics.org/recipe-distribution-basic-boxplot
  
Reference for Violin Plot: 
  - https://r-graphics.org/recipe-distribution-violin#RECIPE-DISTRIBUTION-VIOLIN

- Basic Boxplot

```{r boxplot, echo=TRUE}
ggplot(data, aes(y = faminc_new))+
  geom_boxplot()
```

- Break up boxplots by education group -- add an aesthetic mapping for group

```{r boxplot2, echo=TRUE}
ggplot(data, aes(y = faminc_new, group = educ))+
  geom_boxplot()
```

- Add labels and a title
```{r boxplot3, echo=TRUE}
ggplot(data, aes(y = faminc_new, group = educ))+
  geom_boxplot()+
  labs(title = "Family Inc. by Respondent Ed. Level", x = "Income Level", y = "Family Income")
```

- Reformat the data to create a dichotomous categorical variable for four-year college grads or more, and then all respondents with 2 year college degree or less. 

```{r boxplot4, echo=TRUE}

data$educ_category <- recode(data$educ, '1' = "<4 year Degree", '2' = "<4 year Degree", '3' = "<4 year Degree", '4' = "<4 year Degree", '5' = "4 year Degree", '6' = "4 year Degree +")

# Make sure you change the aesthetic mapping so the new categorical variable is maped to "x" rather than group

ggplot(data, aes(y = faminc_new, x = educ_category))+
  geom_boxplot()+
  labs(title = "Family Inc. by Respondent Ed. Level", x = "Income Level", y = "Family Income")
```

## Histogram (Frequency Distribution Plot)
Reference:
  - https://r-graphics.org/recipe-quick-histogram
  - https://r-graphics.org/recipe-distribution-basic-hist#RECIPE-DISTRIBUTION-BASIC-HIST
  - https://r-graphics.org/recipe-distribution-multi-hist#RECIPE-DISTRIBUTION-MULTI-HIST

```{r boxplot5, echo=TRUE}
ggplot(data, aes(x = faminc_new))+
  geom_histogram()+
  labs(title = "Family Inc. by Respondent Ed. Level", x = "Income Level", y = "Frequency")

# with binwidth
ggplot(data, aes(x = faminc_new))+
  geom_histogram(binwidth = 4)+
  labs(title = "Family Inc. by Respondent Ed. Level", x = "Income Level", y = "Frequency")
```

# Density Plot
Reference: 
- https://r-graphics.org/recipe-distribution-basic-density
```{r boxplot6, echo=TRUE}
ggplot(data, aes(x = faminc_new))+
  geom_density()+
  labs(title = "Family Inc. by Respondent Ed. Level", x = "Income Level", y = "Density")
```