---
title: "Barplot"
output: pdf_document
date: "2022-07-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
```

# Load the Data

```{r dataset}
data <- read_csv("cel_dataset_coursera.csv")
```

# Bar Plot

## Basic Configuration

```{r barplot1}
data %>%
  filter(congress==115)%>%
  ggplot(aes(x = dem))+
  geom_bar()

# prove to yourself your bar plot is right by comparing with a frequency table: 
kable(table(filter(data, congress == 115)$dem))

# use st_name instead, so how counts of how many members of Congress from each state:
data %>%
  filter(congress==115)%>%
  ggplot(aes(x = st_name))+
  geom_bar()

# flip the figure by setting y aesthetic rather than the x
data %>%
  filter(congress==115)%>%
  ggplot(aes(y = st_name))+
  geom_bar()

# let's go back and recode the dem variable to be a categorical variable
party <- recode(data$dem, '1' = "Democrat", '0' = "Republican")

data <- add_column(data, party)

data %>% 
  filter(congress == 115) %>%
  ggplot(aes(x = party))+
  geom_bar()+
  labs(x = "Party", y = "Number of Members")

# Let's manually fill and change the colors of the bars and drop the legends with the guides command
data %>%
  filter(congress == 115) %>%
  ggplot(aes(x = party, fill = party))+
  geom_bar()+
  labs(x = "Party", y = "Number of Members")+
  scale_fill_manual(values = c("blue", "red"))+
  guides(fill = FALSE)
```

## More Configuration
- Making more barplots and manipulating more data in R
### Barplot of proportions
```{r barplot2}
# Create a dummy dataset
apple <- rep("apple", 6)
orange <- rep("orange", 3)
banana <- rep("banana", 1)

fruit_bowl <- tibble("fruits" = c(apple, orange, banana))
fruit_bowl

# create a table that counts fruits in a second column
fruit_bowl_summary <- fruit_bowl %>%
  group_by(fruits) %>%
  summarize("count" = n())

fruit_bowl_summary

# calculate the proportion
fruit_bowl_summary$proportion <- fruit_bowl_summary$count/sum(fruit_bowl_summary$count)

# add the geom_bar, using "stat" to tell command to plot the exact value for proportion
ggplot(fruit_bowl_summary, aes(x = fruits, y = proportion, fill = fruits))+
  geom_bar(stat = "identity")+
  scale_fill_manual(values = c("red", "yellow", "orange"))+
  guides(fill = FALSE)+
  labs(x = "Fruits", y = "Proportion of Fruits")
```

- The `stat = "identity"` in geom_bar means telling the function not to its default action (which is to use that count function) but instead plot the different values for every single row of the dataset directly based on the "identity" of that value in the table.

### Stacked Bar Plot
```{r barplot3}
cces <- read_csv("cces_sample_coursera.csv")

# create counts of Democrats, Republicans, and Independents by Region
cces$region_name <- recode(cces$region, '1' = "Northwest", '2' = "Midwest", '3' = "South", '4' = "West")

cces$dem_rep <- recode(cces$pid7, '1' = "Democrat", '2' = "Democrat", '3' = "Democrat", '4' = "Independent", '5' = "Republican", '6' =  "Republican", '7' = "Republican")

ggplot(cces, aes(x = region_name, fill = dem_rep))+
  geom_bar()
```

### Grouped Bar Plot
- To make a grouped bar chart, we need to set `position` parameter to `"dodge"` inside the `geom_bar` function. This will take the stacked bar chart and breaks it apart.
```{r barplot4}
ggplot(cces, aes(x = region_name, fill = dem_rep))+
  geom_bar(position = "dodge")+
  labs(x = "Region", y = "Count")
```