---
title: "Annotation"
output: pdf_document
date: "2022-07-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction
Elegant Graphics for Data Analysis: 
- https://ggplot2-book.org/annotations.html  
- https://ggplot2-book.org/mastery.html  

We will make a scatter plot of lifetime high scores in teteis against cumulative hours playing tetris. 

Normally, we will see positive correlation.

## Basic Scatter Plot

```{r plot1}
library(tidyverse)

kid <- c("Nick", "Jessica", "Justin", "Brandi", "Kelly", "Enrique")
time_spent <- c(40, 35, 25, 20, 10, 5)
high_score <- c(100, 75, 85, 50, 25, 30)

tetris <- tibble(kid, time_spent, high_score)
tetris

ggplot(tetris, aes(x = time_spent, y = high_score))+
  geom_point()

```

## Scatter Plot with Label
- We can add label on top of the points on the plot using `geom_text` function.

```{r plot2}
ggplot(tetris, aes(x = time_spent, y = high_score))+
  geom_point()+
  geom_text(aes(label = kid))
```

- Notice that the text and the point are overlap making it not look attractive. We have two options here: 
```{r plot3}
# remove the geom_point so only the labels that exist
ggplot(tetris, aes(x = time_spent, y = high_score))+
  #geom_point()+
  geom_text(aes(label = kid))

# use parameter `nudge_x` or `nudge_y` inside `geom_text` function to adjust the label position
ggplot(tetris, aes(x = time_spent, y = high_score))+
  geom_point()+
  geom_text(aes(label = kid), nudge_y = 2.5, nudge_x = 1)

# ?geom_text
```

## Scatter Plot with Congress Dataset
- Plot political ideology and how many bills a member passes
  - Note: __dwnom1__ on the x axis, __all_pass__ on the y axis
  
```{r plot4}
# load the dataset
cel <- drop_na(read_csv("cel_dataset_coursera.csv"))

# use piping to feed the data to ggplot
# notice instead of puting label on geom_text, we put it inside aes() function in ggplot. 
# the subsequent functions (geom_point, geom_text) will inherit the values passed to the params inside the original ggplot function.
cel %>% 
  filter(congress == 115) %>%
  ggplot(aes(x = dwnom1, y = all_pass, label = thomas_name))+
  geom_point()+
  geom_text()
```

- Notice that the plot is __really messy__. 
- One way to solve this is by adding text only for a small subset of the member.
- We can specify to only want the `geom_text` to apply to a subset of the data.

```{r plot5}
cel %>% 
  filter(congress == 115) %>%
  ggplot(aes(x = dwnom1, y = all_pass, label = thomas_name))+
  geom_point()+
  # specify here that you only want the geom_text to apply to a subset of the data
  geom_text(data = filter(cel, congress == 115 & all_pass > 8))
```

- Another way to solve this is by using a new package called `ggrepel`. 
  - This package is not included inside tidyverse, but it works well with ggplot and is pretty common to use for data visualization with R.

```{r plot6}
# install.packages("ggrepel")
library(ggrepel)

# ggrepel will use spacey algorithm to push the text apart from each other making it more readable
cel %>% 
  filter(congress == 115) %>%
  ggplot(aes(x = dwnom1, y = all_pass))+
  geom_point()+
  geom_text_repel(data = filter(cel, congress == 115 & all_pass > 8), mapping = aes(x = dwnom1, y = all_pass, label = thomas_name))

# add rectangular highlight and annotation to the plot
cel %>% 
  filter(congress == 115) %>%
  ggplot(aes(x = dwnom1, y = all_pass))+
  geom_point()+
  geom_text_repel(data = filter(cel, congress == 115 & all_pass > 8), mapping = aes(x = dwnom1, y = all_pass, label = thomas_name))+
  annotate("rect", xmin = .05, xmax = .4, ymin = 13, ymax = 15, alpha = .2, fill = "red")+
  annotate("text", x = .55 , y = 14, label = "Most Passed", color = 'red')
```
