---
title: "New Report with kable and qplot"
output: pdf_document
date: "2022-07-17"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
?rmarkdown

library(tidyverse)
library(knitr)

data <- read_csv('cces_sample_coursera.csv')
data <- drop_na(data)
```

# My First Tables

We can use `kable` function combined to generate a table. 
```{r myTable, echo=TRUE}
?kable
kable(table(data$gender), align = "l")
kable(summarize(data, Mean = mean(data$faminc_new), Median = median(data$faminc_new), align = 'l', label= 'Summary Statistics for Family Income'))
```

## My First QPlots

tidyverse includes one of the most powerful visualization library called `ggplot2` that we will learn later. However, for quick visualization, besides using the R base tool, we can use `qplot` which is a quick plotting tool that incorporates some ggplot tools but seems like based visualization tools in R. 

```{r myQplot, echo=TRUE}
data(mtcars)
# geom stands for geometry (basic options: histogram, density, boxplot, point, and line)

qplot(x = mpg, data = mtcars, geom = 'histogram', bins = 5, main = "Histogram Title", xlab = "Miles per Gallon")
qplot(x = mpg, data = mtcars, geom = 'density', main = "Density Plot Title", xlab = "Miles per Gallon")
qplot(x = mpg, data = mtcars, geom = 'boxplot', main = "Box Plot Title", xlab = "Miles per Gallon")
qplot(x = wt, y = mpg, data = mtcars, geom = 'point', main = "Scatter Plot Title", xlab = "Weight", ylab = "Miles per Gallon")
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
